Research log

### 2024-01-10 docs/ExploratoryResults.Rmd
within-subject variability

@@

>>>>

## Variability in number of misses

```{r battleship-num-misses-variability}

E2.sd_num_clicks <- E2.click_df %>%
  filter(!(subj_id %in% E2.lucky_pretenders))%>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(sd_num_clicks=sd(num_clicks)) %>%
  spread(test_part,sd_num_clicks) %>% 
  mutate(diff=nonpretend-pretend);

E2.sd_num_clicks_relative_to_np <- E2.click_df %>%
  filter(!(subj_id %in% E2.lucky_pretenders))%>%
  group_by(subj_id)%>%
  mutate(mean_nonpretend=mean(num_clicks[test_part=='nonpretend']),
         squared_dist_to_mean = (num_clicks-mean_nonpretend)**2)%>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(sd_num_clicks=sqrt(sum(squared_dist_to_mean)/(n()-1))) %>%
  spread(test_part,sd_num_clicks) %>% 
  mutate(diff=nonpretend-pretend);

We quantified the within-subject variability in the number of misses in pretend and non-pretend games. This is a function of how similar games were to each other, within condition and within individual. The standard deviation in non-pretend games was on average `r E2.sd_num_clicks$nonpretend%>%mean()%>%printnum()`, and significantly higher than in pretend games (mean SD = `r E2.sd_num_clicks$pretend%>%mean()%>%printnum()`, `r apa_print(E2.sd_num_clicks$diff%>%t.test())$statistic`). In other words, pretend games were significantly more similar to each other, in number of guesses, than non-pretend games were to each other.

Next, we asked whether pretend games were not only more similar among themselves, but also to the average non-pretend game, than were non-pretend games. We therefore replaced the $\mu_{pretend}$ in the standard deviation formula with $\mu_{non-pretend}$: $\sigma'=\sqrt{\frac{1}{N}\Sigma_{i=1}^{N} (x_{i}-\mu_{non-pretend})^2}$. Strikingly, we find that a typical pretend game was more similar, in number of misses, **to the average non-pretend game**, than was a typical non-pretend game (`r apa_print(E2.sd_num_clicks_relative_to_np$diff%>%t.test())$full_result`).

@@

## Variability in number of misses

```{r hangman-num-misses-variability}

E4.sd_num_misses <- E4.click_df %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(sd_num_misses=sd(num_misses)) %>%
  spread(test_part,sd_num_misses) %>% 
  mutate(diff=nonpretend-pretend);

E4.sd_num_misses_relative_to_np <- E4.click_df %>%
  group_by(subj_id)%>%
  mutate(mean_nonpretend=mean(num_misses[test_part=='nonpretend']),
         squared_dist_to_mean = (num_misses-mean_nonpretend)**2)%>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(sd_num_clicks=sqrt(sum(squared_dist_to_mean)/(n()-1))) %>%
  spread(test_part,sd_num_clicks) %>% 
  mutate(diff=nonpretend-pretend);

E4.sd_num_misses_relative_to_np_controlling_for_word_difficulty <- E4.click_df %>%
  group_by(word)%>%
  mutate(num_misses_corrected=num_misses-mean(num_misses[test_part=='nonpretend']))%>%
  group_by(subj_id)%>%
  mutate(mean_nonpretend=mean(num_misses_corrected[test_part=='nonpretend']),
         squared_dist_to_mean = (num_misses_corrected-mean_nonpretend)**2)%>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(sd_num_clicks=sqrt(sum(squared_dist_to_mean)/(n()-1))) %>%
  spread(test_part,sd_num_clicks) %>% 
  mutate(diff=nonpretend-pretend);

```
Similar to Battleship, the within-subject standard deviation in number of misses in non-pretend games (`r E4.sd_num_misses$nonpretend%>%mean()%>%printnum()`) was on average significantly higher than in pretend games (mean SD = `r E4.sd_num_misses$pretend%>%mean()%>%printnum()`, `r apa_print(E4.sd_num_misses$diff%>%t.test())$statistic`). Here too, pretend games were significantly more similar to each other, in terms of number of misses, than non-pretend games.

In order to test whether pretend games were more similar to the average non-pretend game than non-pretend games themselves, we first regressed out the effect of word on number of misses, since within-subject differences between the conditions can reflect the random assignment of words to conditions. We then replaced the $\mu_{pretend}$ in the standard deviation formula with $\mu_{non-pretend}$: $\sigma'=\sqrt{\frac{1}{N}\Sigma_{i=1}^{N} (x_{i}-\mu_{non-pretend})^2}$, as we did for Battleship. We again find that a typical pretend game was more similar, in number of misses (controlling for word difficulty), to the average pretend game, than was a typical non-pretend game (`r apa_print(E4.sd_num_misses_relative_to_np_controlling_for_word_difficulty$diff%>%t.test())$full_result`).
